---
title: "`r paste(params$model_run, '- Quick-look Diagnostics')`"
output: 
  html_document:
    toc: true
    toc_float: true
    depth: 2
params:
  model_run: 100
  model_dir: "model/pk"
  log_dv: FALSE
  excludeETAs: !r c() 
  contCov: !r c() # continuous covariates to plot against ETAs
  script: "pk-diagnostics-quick.Rmd"
---

# Purpose

The main purpose of this template is to produce a **set of quick diagnostic 
plots** you can use during the early stages of model development to get a quick 
look at how your model fits the data. 

This template will **create an HTML preview** of some typical diagnostic plots.

These plots are **not** intended to be report quality - please use the
**`pk-diagnostics-report.Rmd`** template to create diagnostics suitable for your
report.


```{r setup_general, include = FALSE, message=FALSE, warning = FALSE}

### General script set-up
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.height = 6, fig.width = 6)

### Libraries ----------------------------
library(tidyverse) 
library(pmplots)
library(bbr)
library(mrgmisc)
library(here) 
library(glue)
library(patchwork)
library(knitr)
library(pmtables)

### Options ----------------------------
options(mrg.script = params$script)
margin <- theme(plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "in"))
parens <-  pmtables::ensure_parens
```


```{r load_any_functions, include = FALSE}
### Source function scripts
source(here('script/functions/functions-diagnostics.R'))
```


```{r setup_bbr, echo = FALSE, message = FALSE}
### Model location 
# Define `modelName` and path to this model directory

# Set directories for bbr models ----------------------------
modelName <- params$model_run
thisModelPath <- file.path(params$model_dir, modelName)
```


# Model details

## Output summaries from bbr

```{r modelOutputSummary, echo = FALSE, results = "asis"}
summ <- mod %>% model_summary()
summ
```

## Annotation and metadata from bbr

```{r modelDetails, echo = FALSE, results = "asis"}
mod <- read_model(here(thisModelPath))
mod
```


```{r read_in_data, echo = FALSE, message=FALSE}

# Use `bbr::nm_join()` to read in the input data set and join to any tables that
# were written out by the model. **This assumes that a row number column (called
# `NUM`) was included during data assembly.** The idea here is that in NONMEM,
# you table just `NUM` and none of the other input data items. They all will get
# joined back to the nonmem output ... even character columns.
data0 <- nm_join(mod)

# filter to observation rows only
data <- data0 %>% 
  filter(EVID==0) 

# If log transformed DV used
if(isTRUE(params$log_dv)) {
  if(all(c("DV", "IPRED", "PRED") %in% names(data0))) {
    data <- mutate(data,
                   DV = exp(DV),
                   IPRED = exp(IPRED),
                   PRED = exp(PRED))  
  } else {
    stop("To use log_dv=TRUE you need to table out DV, PRED, IPRED in your NONMEM model.")
  }
}

# check for npdes - if none then plots will not be created
if("NPDE" %in% names(data)) {
  plotNPDE = TRUE
} else { plotNPDE = FALSE}

# identify the etas included in the dataset 
etas = str_subset(names(data), "^(?i)ETA?")

# the user can choose to exclude specific etas from eta plots, e.g., IOV etas
if(!is_null(params$excludeETAs)){
  etas = etas[which(!etas %in%  params$excludeETAs)]
} 

# check for covariates to plot against ETAs
plotContCov <- length(params$contCov) > 0

# filter to single row per ID and include factors 
id <- distinct(data, ID, .keep_all=TRUE) 
```


# General diagnostic plots

```{r xAxis_label_names, echo = FALSE}
# The following plots assume that the preferred x-axis labels are defined here. 
xTIME <- pm_axis_time()
xTAD <- pm_axis_tad()
xPRED <- pm_axis_pred()
```

## DV vs PRED and IPRED 

Plots of DV vs PRED and IPRED for the full dataset 

```{r DV_PRED_plots, echo = FALSE, message = FALSE}

dvp <- dv_pred(data) + margin      # DV vs PRED 
dvip <- dv_ipred(data) + margin    # DV vs IPRED 

p <- dvp / dvip
p
rm(p, dvp, dvip)
```

## CWRES vs PRED, time and time after dose

```{r cwres_plots, echo = FALSE, message = FALSE}
p <- cwres_panel(data)
p
rm(p)
```

## NPDE plots

NPDE vs PRED, time and time after dose.

```{r NPDE_plots, eval=plotNPDE, echo = FALSE, message=FALSE}
p <- npde_panel(data)
p
rm(p)
```


# EBEs-based diagnostics

## ETA pairs plot 

```{r eta_pairs_plots, echo = FALSE, message = FALSE}
p <- eta_pairs(id, etas)
p
rm(p)
```


```{r eta_cont_cov_title, eval = plotContCov, echo = FALSE, results = "asis"}
cat("## ETA vs continous covariates")
```

```{r eta_cont_cov, eval = plotContCov, echo = FALSE, message = FALSE}

# Create plots with a single ETA and multiple covariates
p <- eta_covariate(id, x = params$contCov, y = etas)
p <- map(p, ~ .x + plot_layout(axes = "collect"))
p
rm(p)
```
